---
title: "Holsenbeck_S_5"
author: "Stephen Synchronicity"
date: "2017-10-09"
output:
  html_document:
    css: C:\Users\Stephen\Documents\R\win-library\3.4\rmarkdown\rmarkdown\templates\neu_hwk\resources\styles.css
    highlight: zenburn
    keep_md: yes
    theme: readable
    toc: yes
    toc_float: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=F,cache=TRUE, fig.align='center', fig.height=3.5, fig.width=5, tidy=TRUE, tidy.opts=list(width.cutoff=80))
rmarkdown::html_dependency_jquery()
rmarkdown::html_dependency_bootstrap("readable")
rmarkdown::html_dependency_jqueryui()
set.seed(1)
```
```{r Libraries, echo=FALSE, results='hide'}
library("tidyverse")
library(lubridate)
library("htmltools")
library("DT")
library("stringr")
```

#Homework 5
<em style="color:red;"><strong>Note:</strong> The last R chunk for section 4 uses 4 mapply functions with control loops that can take up to 10 minutes to compute. The chunk options from this chunk on are set to 'eval=F' to prevent this from running until manually done so to prevent any confusion if all chunks were run and R computed for 10+ mins. </em>

## 1
(10 points) Add a new column `Season1Days` that contains the number of days a market is opened per week (for the dates it is open).
```{r 'Load Data'}
fmdb <- as.tibble(read_csv(file = "C:\\Users\\Stephen\\Documents\\Northeastern\\DA 5020 - Collect Store Retrieve Data\\Homework\\HWK4\\farmers_market.csv"))
```
```{r '1'}
# fmdb <- fmdb %>%
#   separate(col="Season1Date",into=c("S1B","S1E"),sep="to",remove=F)
# m2n<- function (x){match(tolower(x), tolower(month.name))}
# m2n("January")
#Track whether it's open or closed
fmdb$Season1Days <- str_count(fmdb$Season1Time, "[[:alpha:]]{3}")
  #separate(col="Season1Time",into=c("S1D","S1T"),sep=":",remove=F,extra="merge")
```
## 2
(10 points) Add a new column `WeekendOpen` indicating whether a market opens during weekends in `Season1`
```{r '2'}
fmdb$WeekendOpen <- str_replace(fmdb$Season1Time,"(?:.*)((?:Sat:[\\s:0-9AaPpMm-]+;)+)(?:.*)|(?:.*)((?:Sun:[\\s:0-9AaPpMm-]+;)+)(?:.*)|.*","\\1\\2")
```
## 3
(20 points) Find out which markets close before 6PM, and which open only for fewer than 4 hours a day. For simplicity, consider only `Season1Time`. For markets with different open hours across a week, use the average length of open hours for the days they actually open.
```{r '3: Days Table'}
#Markets that close before 6p
(fmtimes <- fmdb %>% #create dataframe with individual day times
   select(FMID,Season1Time) %>% 
   mutate(Mon = str_extract(fmdb$Season1Time,"((?:Mon:[\\s:0-9AaPpMm-]+;)+)"),
          Tue = str_extract(fmdb$Season1Time,"((?:Tue:[\\s:0-9AaPpMm-]+;)+)"),
          Wed = str_extract(fmdb$Season1Time,"((?:Wed:[\\s:0-9AaPpMm-]+;)+)"),
          Thu = str_extract(fmdb$Season1Time,"((?:Thu:[\\s:0-9AaPpMm-]+;)+)"),
          Fri = str_extract(fmdb$Season1Time,"((?:Fri:[\\s:0-9AaPpMm-]+;)+)"),
          Sat = str_extract(fmdb$Season1Time,"((?:Sat:[\\s:0-9AaPpMm-]+;)+)"),
          Sun = str_extract(fmdb$Season1Time,"((?:Sun:[\\s:0-9AaPpMm-]+;)+)"))
)
```
```{r '3 Before/After 6 Table'}
fmsix <- fmtimes %>% 
    mutate('After6' =str_extract(Season1Time,pattern=regex("((?:[7-9]).{2,3}\\s[PpMm]{2})")))
filter(`fmsix`,is.na(fmsix$After6)) #filter by those without times, ie close before 6 for verification purposes

```
```{r '3: Open for <4Hrs', results='hide'}
wd <- data.frame(c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"),"sta","end",stringsAsFactors = F)
wd$Start  <- do.call(paste, c(wd[1],wd[2],sep = "."))
wd$End  <- do.call(paste, c(wd[1],wd[3],sep = "."))#Vector for naming

fmsix <- fmsix %>% #Sepearate the times into cols
  separate(col=Mon,into=c("Mon","Times"),sep=":",extra="merge") %>% 
  separate(col=Times,into=c(wd$Start[1],wd$End[1]),sep="-",extra="merge") %>% 
   separate(col=Tue,into=c("Tue","Times"),sep=":",extra="merge") %>% 
  separate(col=Times,into=c(wd$Start[2],wd$End[2]),sep="-",extra="merge") %>%
   separate(col=Wed,into=c("Wed","Times"),sep=":",extra="merge") %>% 
  separate(col=Times,into=c(wd$Start[3],wd$End[3]),sep="-",extra="merge") %>%
   separate(col=Thu,into=c("Thu","Times"),sep=":",extra="merge") %>% 
  separate(col=Times,into=c(wd$Start[4],wd$End[4]),sep="-",extra="merge") %>%
   separate(col=Fri,into=c("Fri","Times"),sep=":",extra="merge") %>% 
  separate(col=Times,into=c(wd$Start[5],wd$End[5]),sep="-",extra="merge") %>%
   separate(col=Sat,into=c("Sat","Times"),sep=":",extra="merge") %>% 
  separate(col=Times,into=c(wd$Start[6],wd$End[6]),sep="-",extra="merge") %>%
   separate(col=Sun,into=c("Sun","Times"),sep=":",extra="merge") %>% 
  separate(col=Times,into=c(wd$Start[7],wd$End[7]),sep="-",extra="merge")
tCols <- grep("(.*.sta)|(.*.end)",names(fmsix)) #Vector of cols needing time transforms
#strptime(fmsix$Tue.sta[5],format="%H:%M") Test Function
fmsix <- cbind(fmsix,as.data.frame(apply(fmsix[,tCols], 2, function(x) strptime(x, "%I:%M %p")))) #Convert time strings to posix times, cbind them in 24 Hr format
fmsix <- subset(fmsix, select=-seq(from=(min(tCols)-1),to=max(tCols))) #remove previous time string cols
fmsix <- fmsix %>% #Add cols w/ numeric # of hrs difference
  mutate(Mon.hrs = as.numeric(abs(difftime(Mon.sta,Mon.end,units="hours"))),
         Tue.hrs = as.numeric(abs(difftime(Tue.sta,Tue.end,units="hours"))),
         Wed.hrs = as.numeric(abs(difftime(Wed.sta,Wed.end,units="hours"))),
         Thu.hrs = as.numeric(abs(difftime(Thu.sta,Thu.end,units="hours"))),
         Fri.hrs = as.numeric(abs(difftime(Fri.sta,Fri.end,units="hours"))),
         Sat.hrs = as.numeric(abs(difftime(Sat.sta,Sat.end,units="hours"))),
         Sun.hrs = as.numeric(abs(difftime(Sun.sta,Sun.end,units="hours")))) 
tCols <- grep("(.*.hrs)",names(fmsix)) #refresh the cols vector
fmsix <- transform(fmsix, mHrs = rowMeans(fmsix[,tCols], na.rm = TRUE)) #add a col with means
fmsix %>% 
  filter(mHrs<4) #filter for those less than 4

#tCols <- grep("(.*.sta)|(.*.end)",names(fmsix)) #refresh the cols vector
#eCols <- seq(from=min(tCols),to=max(tCols),by=2)
# test <- for(i in eCols){
#   print(i)
#   as.data.frame(apply(fmsix[,eCols],2,function(i) difftime(fmsix[,i],fmsix[,(i+1)],units="hours")))
# }
```

## 4
(40 Points) The seasons are not standardized and would make analysis difficult. Create four new columns for four seasons (Spring, Summer, Fall, Winter), indicating whether a market is available in that season. Also, create two additional columns `HalfYear` and `YearRound` to identify those who open across seasons. Define "half year" and "year round" on your own terms, but explain them before you write the code (or as comments in your code). (Hint: you may want to create even more auxiliary columns, `Season1BeginDate` and `Season1EndDate` for example.)
```{r '4: Seasons Dataframe'}
fmSeasons <- fmdb %>%
  select(FMID,Season1Date,Season2Date,Season3Date,Season4Date) %>% 
  separate(Season1Date,into=c("S1.beg","S1.end"),sep=" to ",extra="merge") %>%
  separate(Season2Date,into=c("S2.beg","S2.end"),sep=" to ",extra="merge") %>%
  separate(Season3Date,into=c("S3.beg","S3.end"),sep=" to ",extra="merge") %>%
  separate(Season4Date,into=c("S4.beg","S4.end"),sep=" to ",extra="merge")
#month.num<- seq(as.Date("2017/1/1"), by = "month", length.out = 12)
```
```{r '4: String Xforms'}
tCols <- grep("(S\\d.beg)|(S\\d.end)",names(fmSeasons))#refresh the cols vector
# month.namestonums <- function(x) {
#   str_replace_all(x,"January","1")
#   str_replace_all(x,"February","2")
#   str_replace_all(x,"March","3")
#   str_replace_all(x,"April","4")
#   str_replace_all(x,"May","5")
#   str_replace_all(x,"June","6")
#   str_replace_all(x,"July","7")
#   str_replace_all(x,"August","8")
#   str_replace_all(x,"September","9")
#   str_replace_all(x,"October","10")
#   str_replace_all(x,"November","11")
#   str_replace_all(x,"December","12")
# }
# fmSeasons <- apply(fmSeasons,2,function(x){month.namestonums(x)})

fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"January","1")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"February","2")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"March","3")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"April","4")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"May","5")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"June","6")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"July","7")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"August","8")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"September","9")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"October","10")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"November","11")})
fmSeasons <- apply(fmSeasons,2,function(x){str_replace_all(x,"December","12")})
fmSeasons <- apply(fmSeasons,2,function(x){gsub("(^\\d{1,2}\\b(?!\\/))","\\1\\/1\\/2017",perl=T,x)})# Make months into date string
fmSeasons <- apply(fmSeasons,2,function(x){gsub("(\\d{1,2}\\/\\d{1,2}\\/\\d{2,4}).*","\\1",perl=T,x)}) #Make sure all are mdy dates
fmSeasons <- as.tibble(fmSeasons) 
fmSeasons <- fmSeasons  %>%  
  mutate_at(vars(matches("(beg)|(end)")), funs(mdy)) %>% 
  mutate(
    s1.int = S1.beg %--% S1.end,
    s2.int = S2.beg %--% S2.end,
    s3.int = S3.beg %--% S3.end,
    s4.int = S4.beg %--% S4.end)
# Table with dates and intervals
fmInts <- fmSeasons %>% 
  mutate_at(vars(s1.int:s4.int),funs(as.period(.)))#
#With intervals as periods
```
```{r '4: Year Rounds'}
#all markets longer than 11mo
tCols <- grep("s\\d.int",names(fmSeasons))#refresh the cols vector
fmYR <- fmSeasons %>% 
  mutate_at(vars(s1.int:s4.int),funs(as.period(.)))%>%
  mutate(yearRound = ifelse(int_length(fmSeasons$s1.int)>as.period(months(11),unit="seconds"),T,F)) 
filter(fmYR,yearRound == T)

```
```{r '4: Half Year'}
#All Markets longer than 6mo and less than 11 mo
fmHY <- fmInts %>% 
  mutate(halfYear = ifelse(!is.na(fmInts$s1.int) & fmInts$s1.int<as.period(months(11)) & fmInts$s1.int>as.period(months(6)),T,NA)) 
filter(fmHY,halfYear == T)
rm(fmInts)
```
```{r '4: Join the tables'}
fmSeasonsDetail <- left_join(fmSeasons,fmYR[,c(1,14)],by='FMID')
fmSeasonsDetail <- left_join(fmSeasonsDetail,fmHY[,c(1,14)],by='FMID')
```
```{r '4: Seasonal Ranges'}
#Testing Int_overlaps, returns logical 
#int_overlaps(fmSeasonsDetail$s1.int[[1]],interval(ymd('2017-2-1'),ymd('2017-7-1')))
seasons<- tibble::tribble(
      ~Date,
        "March 20",
       "June 21",
     "September 22",
   "December 21")# Seasons Table
seasons <- mdy(paste(seasons$Date,year(today()))) # makes dates
spring <- interval(seasons[1],seasons[2])
summer <- interval(seasons[2],seasons[3])
fall <- interval(seasons[3],seasons[4])
winter <- interval(seasons[4]-years(1),seasons[1]) # define each as intervals
#Season overlap functions
#Input: Market interval
#Function: defines seasonal intervals based on yr of input interval
#Output: Returns overlap of market interval and season interval
sOlp.sp <- function(x){
  if(!is.na(x)){
  y <- year(int_start(x))
  year(seasons) <- y
  sn <- seasons
  sp <- interval(sn[1],sn[2])
  return(intersect(x,sp))
  }else{
    return(NA)
  }
}
sOlp.su <- function(x){
  if(!is.na(x)){
  y <- year(int_start(x))
  year(seasons) <- y
  sn <- seasons
  su <- interval(sn[2],sn[3])
  return(intersect(x,su))}else{return(NA)}
}
sOlp.fa <- function(x){
  if(!is.na(x)){
  y <- year(int_start(x))
  year(seasons) <- y
  sn <- seasons
  fa <- interval(sn[3],sn[4])
  return(intersect(x,fa))}else{return(NA)}
}
sOlp.wi <- function(x){
  if(!is.na(x)){
  y <- year(int_start(x))
  year(seasons) <- y
  sn <- seasons
  year(sn[1]) <- y+1
  wi <- interval(sn[4],sn[1])
  return(intersect(x,wi)) #return overlap
  }else{return(NA)}
}
# test <- test %>%
#   group_by(s1.int) %>% 
#   filter(!is.na(S1.end))%>%  does not work to filter table, fn(x)s updated to # handle Na
#   mutate(Spring=sOlp.sp(s1.int))
#is.na(test$s1.int[[3]])
#sOlp.sp(test$s1.int[[1]]) #Testing the Functions
```
<div class="a">It's worth noting here that the winter season presents quite a bit of complexity when attempting to find the overlap between the winter season and the market intervals. If I attempted to continue with the same inquiry of determining the overlap interval with each season, there would need to an iterative loop for winter, that accounted for five different cases, which are:
1. The beginning of the market inverval overlaps with winter.
2. The end of the market inverval overlaps with winter.
3. The beginning and end of the market inverval overlaps with winter, with no gap (year round).
4. The beginning and end of the market inverval overlaps with winter, with a gap (during December/the holidays).
5. The market interval does not overlap with winter.
Due to time constraints, rather than get mired in writing this loop and then having to mapply it (which will likely take forever) I'm going to stick with the same method used for the previous seasons. While imperfect, it will have to do in this instance. 
</div>
```{r '4: Create Season Dataframes', eval=F,results='hide'}
#Create seasonal interval dataframes
test <- fmSeasonsDetail
Spring <- as.data.frame(do.call(c,mapply(sOlp.sp, test$s1.int,SIMPLIFY = F)))
colnames(Spring) <- "Spring"
Summer <- as.data.frame(do.call(c,mapply(sOlp.su, test$s1.int,SIMPLIFY = F)))
colnames(Summer) <- "Summer"
Fall <- as.data.frame(do.call(c,mapply(sOlp.fa, test$s1.int,SIMPLIFY = F)))
colnames(Fall) <- "Fall"
Winter <- as.data.frame(do.call(c,mapply(sOlp.wi, test$s1.int,SIMPLIFY = F)))
colnames(Winter) <- "Winter"
```
```{r 'Unify Market time data',eval=T,results='hide'}
fmSeasonsDetail <- cbind(fmSeasonsDetail,Spring,Summer,Fall,Winter) #View all in single data frame
rm(Spring,Summer,Fall,Winter,test,fmSeasons,fmHY,fmYR)#Free up Space
fmSeasonsDetail$FMID <- as.numeric(fmSeasonsDetail$FMID)
fmdb <- left_join(fmdb,fmSeasonsDetail[,c(1,14:19)],by='FMID')
```

## 5 
(20 points) _Open question_: explore the new variables you just created. Aggregate them at different geographic levels, or some other categorical variable. What can you discover?

```{r '5',eval=F}
#prereqs
# install.packages(c("maps", "mapdata")) 
# devtools::install_github("dkahle/ggmap")
library(maps)
library(mapdata)
library(ggplot2)
library(ggmap)
states <- map_data("state")
usmap <- ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = F, group = group), fill = NA, color = "red") +
  coord_fixed(1.3) +
  guides(fill=FALSE)
cont.usa <- fmdb%>%
  filter(State!="Alaska"&State!="Hawaii")
usmap +
geom_point(data=cont.usa,mapping=aes(x=x,y=y,color=yearRound))
#Source:http://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html
```
<div class="a">A)With few exceptions, the year round farmer's markets are located primarily in the southern territories including Puerto Rico and the US Virgin Islands. <em>Note: Alaska and Hawaii not included due to the distortion of scale.</em></div>
<hr>




Below is a test module that was used when lubridate was malfunctioning. A restart ameliorated the issue. I kept the code in case a bug report needed to be made. 
```{r 'Winter Overlap Error', eval=F}
seasons<- tibble::tribble(
      ~Date,
        "March 20",
       "June 21",
     "September 22",
   "December 21")# Seasons Table
seasons <- mdy(paste(seasons$Date,year(today()))) # makes dates
spring <- interval(seasons[1],seasons[2])
summer <- interval(seasons[2],seasons[3])
fall <- interval(seasons[3],seasons[4])
winter <- interval(seasons[4]-years(1),seasons[1]) # define each as intervals

sOlp.wi <- function(x){
  if(!is.na(x)){
  y <- year(int_start(x))
  year(seasons) <- y
  sn <- seasons
  year(sn[1]) <- y+1
  wi <- interval(sn[4],sn[1])
  return(intersect(x,wi)) #return overlap
  }else{return(NA)}
}
fmSeasons$s1.int[[2870]]
(test <- interval("2017-01-01","2017-12-31"))
sOlp.wi(fmSeasons$s1.int[[2870]])

```

